\chapter{Concepts} \label{sec:Concepts}

\todoall{Description of elements should follow this structure: What is it for? What does it do? Examples in concrete syntax + explanation}

\section*{Old stuff from rejected paper}
In this section, we will briefly introduce story diagrams and their current features. 
Story diagrams combine
UML activity diagrams and graph transformations by embedding graph replacement rules into the activities.
This allows the activities in Figure \ref{fig:transformationOverview} to be specified formally by graph replacements while preserving the general control flow structure of the example transformation.

In terms of the classification of model transformations proposed by Czarnecki and Helsen~\cite{Czarnecki06}, story diagrams are an endogenous, in-place transformation language.
It has both declarative (pattern matching) and operational elements (specification of control flow):
Its control flow allows a deterministic selection of the graph replacement rules to be applied, with a (non-deterministic) pattern matching in the graph replacement rules.
It can also be used for inter-model transformations to create a new target model from a given source model, as seen in the example given in this paper. 
In order to execute story diagrams, code generation \cite{GBD07} and interpretation \cite{GHS09} are supported.

In the following, we will describe the graph transformations, the so-called story patterns, in Section~\ref{sec:StoryPatterns}.
Afterwards, we will explain how control flow can be modeled by using elements from activity diagrams in Section~\ref{sec:StoryDiagrams}.

	\input{02b_Concept_StoryPattern.tex} 
	
	\input{02c_Concept_StoryDiagrams.tex}

	\input{02a_Concept_Expressions.tex}
	
	\input{Concept_Templates.tex}
	
 